---
title: "Human Activity Recognition Using Smartphones Dataset"
author: "Gambera Matteo"
date: "05-05-2020"
output:
  pdf_document: default



--- 
The purpose of this project is to demonstrate ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis.
---
1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement.
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names.
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

# Load Data
```{r}

library(reshape2)
packages <- c("data.table", "reshape2")
sapply(packages, require, character.only=TRUE, quietly=TRUE)
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
# i can use also getwd() for directory path
path <- getwd()
#path <- "/home/matteo/Scrivania/datasciencecoursera/Course_2/Data"
# download the zip file
download.file(url, file.path(path, "dataFiles.zip"))
# unzip
unzip(zipfile = "dataFiles.zip")

```
*fread*
_Similar to read.table but faster and more convenient. All controls such as sep, colClasses and nrows are automatically detected. bit64::integer64 types are also detected and read directly without needing to read as character before converting._

# Clean Data
```{r}
#check the name of what type of activity the accelerometer registers
activityLabels <- fread(file.path(path, "UCI HAR Dataset/activity_labels.txt"), col.names = c("Labels", "Activity"))
#view
activityLabels
str(activityLabels)
```
_There are 6 type of activity_
```{r}
features <- fread(file.path(path, "UCI HAR Dataset/features.txt"), col.names = c("id", "featureNames"))
head(features)
str(features) 
```
_I don't want all features(exp ENTROPY),i select and use the only ones i'm interested (words with mean and std)_
```{r}
featuresid <- grep("(mean|std)\\(\\)", features[, featureNames])
featuresid
```
_check for correct grep_
```{r}
meanstdmeasure <- features[featuresid, featureNames]
meanstdmeasure 

```
_cut "()" and  from feauturesNames_
```{r}
meanstdmeasure <- gsub('[()]', '', meanstdmeasure)
meanstdmeasure
```
# Load Train
_load the train dataset_
```{r}
train <- fread(file.path(path, "UCI HAR Dataset/train/X_train.txt"))[, featuresid, with = FALSE]
data.table::setnames(train, colnames(train), meanstdmeasure) #rename the columns with meanstdmeasure 
head(train, n=1)
str(train)
```
_load train activities and subjects_
```{r}
trainActivities <- fread(file.path(path, "UCI HAR Dataset/train/y_train.txt"), col.names = c("Activity"))
trainActivities

trainSubjects <- fread(file.path(path, "UCI HAR Dataset/train/subject_train.txt"), col.names = c("SubjectNum"))
trainSubjects
```
_merge this data to train_
```{r}
train <- cbind(trainActivities,trainSubjects,train)
head(train, n=1)
```
# Load test
_same as train_
```{r}
test <- fread(file.path(path, "UCI HAR Dataset/test/X_test.txt"))[, featuresid, with = FALSE]
data.table::setnames(test, colnames(test), meanstdmeasure)
testActivities <- fread(file.path(path, "UCI HAR Dataset/test/y_test.txt")
                        , col.names = c("Activity"))
testSubjects <- fread(file.path(path, "UCI HAR Dataset/test/subject_test.txt")
                      , col.names = c("SubjectNum"))
test <- cbind(testSubjects, testActivities, test)
head(test, n=1)
```

# Merge test and train
```{r}
data <- rbind(train, test)
str(data)
nrow(data)==sum(nrow(test)+nrow(train))
```
```{r}
data[["Activity"]] <- factor(data[, Activity], levels = activityLabels[["classLabels"]], labels = activityLabels[["activityName"]])
data[["SubjectNum"]] <- as.factor(data[, SubjectNum])
data <- reshape2::melt(data = data, id = c("SubjectNum", "Activity"))
data <- reshape2::dcast(data = data, SubjectNum + Activity ~ variable, fun.aggregate = mean)

```
```{r}
data.table::fwrite(x = data, file = "tidyData.txt", quote = FALSE)
```


